---
title: "FinalExam-2802499756-Nysa Setiawan"
author: "Nysa Setiawan"
date: "2025-06-23"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(  
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev = "png",
  fig.width=5,
  fig.height=7
)
```
## Case Studies {.tabset}
Link Video: https://drive.google.com/file/d/16jRZ6qN91Sdu25oLH7q_W2Ewa237gZ2E/view?usp=sharing

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
```
### Identify data quality issues 

```{r}
df<-read.csv("data_D.csv")
head(df)
```
This is the head of the "data_D.csv"

**Missing Values**
```{r}
colSums(is.na(df))
```
It can be seen there are 12 missing values in "duration" column.

**Duplicated Values**
```{r}
sum(duplicated(df))
```
There are no duplicated values in this dataset.

**Inconsistent Categories**
```{r}
lapply(df[, !names(df) %in% "X"], unique)
```
It can be seen there are "" and "unknown" indicating missing value within job, marital, education, default, housing and loan column.<br>
There are noisy value which is "999" in pdays column.

**Outliers**
```{r}
plot_ly(data = df, y = ~duration, type = "box", name = "Duration") %>%
  layout(title = "Before: Boxplot of Duration",
         yaxis = list(title = "Duration"))
plot_ly(data = df, y = ~previous, type = "box", name = "Previous") %>%
  layout(title = "Before: Boxplot of Previous",
         yaxis = list(title = "Previous"))
plot_ly(data = df, y = ~campaign, type = "box", name = "Campaign") %>%
  layout(title = "Before: Boxplot of Campaign",
         yaxis = list(title = "Campaign"))
```
Although there are many outliers that need to be addressed in later steps, the "previous" variable will not be removed, as it is expected to have an influence for "y" column.

**Handle the issues**
```{r}
df <- df %>%
  mutate(
    pdays = as.numeric(pdays),  
    pdays_new = case_when(
      pdays == 999 ~ "not previously contacted",
      pdays < 7    ~ "in a week",
      pdays <= 14  ~ "more than one week",
      pdays > 14   ~ "more than two weeks"
    )
  )
```
1. Convert the "pdays" column into a categorical variable. Since the value 999 appears frequently in the dataset, it should not be removed but instead categorized appropriately.

```{r}
df[df == "unknown"] <- NA
df[df == ""] <- NA
```
2. Change "unknown" and "" into NA.

```{r}
colMeans(is.na(df)) * 100
```
It can be see the percentages of missing value in the categorical value
```{r}
get_mode <- function(x) {
  ux <- unique(x[!is.na(x)])        
  ux[which.max(tabulate(match(x, ux)))]
}
df$job[is.na(df$job)] <- get_mode(df$job)
df$marital[is.na(df$marital)] <- get_mode(df$marital)
df$education[is.na(df$education)] <- get_mode(df$education)
df$default[is.na(df$default)] <- get_mode(df$default)
df$housing[is.na(df$housing)] <- get_mode(df$housing)
df$loan[is.na(df$loan)] <- get_mode(df$loan)
df$duration[is.na(df$duration)] <- median(df$duration, na.rm = TRUE)
```
3. Fill the NA in categorical columns with mode.<br>
4. Fill the NA in numerical columns (duration) with median.


```{r}
numeric_cols <- sapply(df, is.numeric)
exclude_cols <- c("age", "pdays", "previous")
cols_to_winsor <- setdiff(names(df)[numeric_cols], exclude_cols)

for (colname in cols_to_winsor) {
  Q1 <- quantile(df[[colname]], 0.25, na.rm = TRUE)
  Q3 <- quantile(df[[colname]], 0.75, na.rm = TRUE)
  IQR_val <- Q3 - Q1

  lower_bound <- Q1 - 1.5 * IQR_val
  upper_bound <- Q3 + 1.5 * IQR_val

df[[colname]] <- ifelse(df[[colname]] < lower_bound, lower_bound,
                        ifelse(df[[colname]] > upper_bound, upper_bound,
                               df[[colname]]))

}
```
5. Handle the outliers with capping, but exclude age, pdays and previous.

```{r}
plot_ly(data = df, y = ~duration, type = "box", name = "Duration") %>%
  layout(title = "After: Boxplot of Duration",
         yaxis = list(title = "Duration"))
plot_ly(data = df, y = ~campaign, type = "box", name = "Campaign") %>%
  layout(title = "After: Boxplot of Campaign",
         yaxis = list(title = "Campaign"))
```
These are boxplot of "Duration" and "Campaign" after capping.

```{r}
write.csv(df, "clean_data.csv", row.names = FALSE)
clean_data<-df
head(clean_data)
```
5. Create a clean data set "clean_data"

**Summary statistics and initial exploration**

**Central Tendency and Variability**
```{r}
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

num_cols <- sapply(clean_data, is.numeric)
numeric_data <- clean_data[, num_cols]
numeric_data <- numeric_data[, !(names(numeric_data) %in% c("X", "pdays"))]

summary_stats <- data.frame(
  Mean = sapply(numeric_data, mean, na.rm = TRUE),
  Median = sapply(numeric_data, median, na.rm = TRUE),
  Mode = sapply(numeric_data, get_mode),
  SD = sapply(numeric_data, sd, na.rm = TRUE),
  Variance = sapply(numeric_data, var, na.rm = TRUE),
  IQR = sapply(numeric_data, IQR, na.rm = TRUE),
  Min = sapply(numeric_data, min, na.rm = TRUE),
  Max = sapply(numeric_data, max, na.rm = TRUE),
  Range = sapply(numeric_data, function(x) diff(range(x, na.rm = TRUE)))
)

print(summary_stats)

```

**Summary Table**
```{r}
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

summary_by_group <- function(data, num_col, cat_col) {
  aggregated <- aggregate(
    data[[num_col]] ~ data[[cat_col]],
    data = data,
    FUN = function(x) c(
      mean = mean(x, na.rm = TRUE),
      median = median(x, na.rm = TRUE),
      mode = get_mode(x),
      sd = sd(x, na.rm = TRUE),
      var = var(x, na.rm = TRUE),
      iqr = IQR(x, na.rm = TRUE),
      min = min(x, na.rm = TRUE),
      max = max(x, na.rm = TRUE),
      range = diff(range(x, na.rm = TRUE))
    )
  )

  result <- data.frame(
    Group = aggregated[[1]],
    Mean = aggregated[[2]][, "mean"],
    Median = aggregated[[2]][, "median"],
    Mode = aggregated[[2]][, "mode"],
    SD = aggregated[[2]][, "sd"],
    Variance = aggregated[[2]][, "var"],
    IQR = aggregated[[2]][, "iqr"],
    Min = aggregated[[2]][, "min"],
    Max = aggregated[[2]][, "max"],
    Range = aggregated[[2]][, "range"]
  )

  names(result)[1] <- cat_col
  return(result)
}

summary_age_by_job <- summary_by_group(df, "age", "job")
summary_age_by_marital <- summary_by_group(df, "age", "marital")
summary_age_by_education <- summary_by_group(df, "age", "education")

summary_age_by_default <- summary_by_group(df, "age", "default")
summary_age_by_housing <- summary_by_group(df, "age", "housing")
summary_age_by_loan <- summary_by_group(df, "age", "loan")

summary_age_by_contact <- summary_by_group(df, "age", "contact")
summary_age_by_month <- summary_by_group(df, "age", "month")
summary_age_by_day_of_week <- summary_by_group(df, "age", "day_of_week")

summary_age_by_poutcome <- summary_by_group(df, "age", "poutcome")
summary_age_by_y <- summary_by_group(df, "age", "y")
summary_age_by_pdays_new <- summary_by_group(df, "age", "pdays_new")

print(summary_age_by_job)
print(summary_age_by_marital)
print(summary_age_by_education)
print(summary_age_by_default)
print(summary_age_by_housing)
print(summary_age_by_loan)
print(summary_age_by_contact)
print(summary_age_by_month)
print(summary_age_by_day_of_week)
print(summary_age_by_poutcome)
print(summary_age_by_y)
print(summary_age_by_pdays_new)
```
These are summary table between age with 12 categorical columns. 

### Relationship Analysis
**Explore Variable Relationship**
```{r}
tab_job_y <- table(clean_data$job, clean_data$y)
addmargins(tab_job_y)

prop_total <- prop.table(tab_job_y) * 100
prop_percentage_job_y <- addmargins(round(prop_total, 2))
print(prop_percentage_job_y)

conversion_rate <- prop.table(tab_job_y, margin = 1) * 100
conversion_rate_rounded <- round(conversion_rate, 2)
print(conversion_rate_rounded)
```
From the distribution of job types to campaign outcomes, several insights can be drawn. The vast majority of clients, nearly 89%, did not subscribe to the term deposit, highlighting the overall low success rate of the campaign. Among the various job groups, admin. and blue-collar workers make up the largest portions of the dataset, contributing 22.38% and 21.26% respectively to the “no” group, and 3.18% and 1.44% respectively to the “yes” group. While this makes them the top contributors to the total number of subscribers in absolute terms, it is largely due to their high representation in the client base rather than their actual responsiveness to the campaign.<br>

Based on the data, technician clients represent the second-highest contributor to the number of successful campaign outcomes, accounting for 1.82% of all subscriptions. This places them just behind administrative workers, who lead with 3.18%.<br>

On the other hand, certain job groups such as students and retired individuals, though smaller in number, show disproportionately higher representation in the “yes” category compared to their share in the overall dataset. For example, students contribute only 1.58% to the “no” group but 0.66% to the “yes” group, while retirees account for 3.35% of “no” responses but 1.09% of “yes” responses. This suggests a higher conversion or success rate within these groups. Student and retirees showing a better conversion rate of 24.59% and 29.35%.<br>

```{r}
tab_marital_y <- table(clean_data$marital, clean_data$y)
addmargins(tab_marital_y)

prop_total <- prop.table(tab_marital_y) * 100
prop_percentage_marital_y <- addmargins(round(prop_total, 2))
print(prop_percentage_marital_y)

conversion_rate <- prop.table(tab_marital_y, margin = 1) * 100
conversion_rate_rounded <- round(conversion_rate, 2)
print(conversion_rate_rounded)

```
Based on the data, most of the clients are married, making up around 61% of the total. They also make up the biggest portion of both the people who didn’t subscribe and those who did—over 54% of the “no” responses and about 6% of the “yes” responses. This shows that while many married clients were contacted, only a small portion of them actually subscribed, meaning their success rate is relatively low compared to how many there are.<br>

Single clients, who make up about 28% of the data, contributed nearly 4% of the total subscriptions. Meanwhile, divorced clients form the smallest group, with only around 11% of the total and just over 1% of the “yes” responses. Even though married clients have the most subscriptions in total, that’s mostly because there are so many of them. In contrast, single clients, though fewer in number, seem to be more likely to subscribe when contacted—suggesting they might be more open or responsive to the campaign.<br>

In summary, marital status seems to influence how clients respond, with single clients showing a better conversion rate of 13.17% than married or divorced ones.<br>

```{r}
tab_education_y <- table(clean_data$education, clean_data$y)
addmargins(tab_education_y)

prop_total <- prop.table(tab_education_y) * 100
prop_percentage_education_y <- addmargins(round(prop_total, 2))
print(prop_percentage_education_y)

tab_edu_y <- table(clean_data$education, clean_data$y)
conversion_rate <- prop.table(tab_edu_y, margin = 1) * 100
conversion_rate_rounded <- round(conversion_rate, 2)
print(conversion_rate_rounded)
```
Based on the data, clients with a university degree form the largest segment in the dataset, accounting for 33.11% of all clients and 4.38% of all term deposit subscriptions. This makes them the biggest contributor to the overall number of successful campaign outcomes. The second-largest group is high school graduates, representing 23.24% of the dataset and 2.80% of total subscriptions. Clients with professional courses also hold a notable share at 12.66%, followed by those with basic education levels (4y, 6y, 9y), which together account for about 30.9% of the dataset. Although these basic education groups are sizable, their contributions to the total number of “yes” responses are relatively small. The illiterate group is extremely small, making up only 0.07% of the data, with just 0.01% of total subscriptions.<br>

University graduates have a strong conversion rate of 13.23%, indicating that a sizable portion of them subscribed when contacted. High school graduates also show good responsiveness with a conversion rate of 12.06%. Meanwhile, those with professional course backgrounds convert at 9.88%, and basic education groups fall even lower, particularly basic.9y, which has the lowest rate at 6.76%. On the other hand, the illiterate group which have a small sample in the dataset shows the highest conversion rate at 16.67%.<br>

**Feature Engineering**
```{r}
clean_data$age_group <- ifelse(clean_data$age < 30, "<30",
                 ifelse(clean_data$age <= 50, "30–50", ">50"))
clean_data$recency_level <- ifelse(clean_data$pdays == 999, "never",
                     ifelse(clean_data$pdays <= 7, "recent", "old"))
clean_data$total_contact_score <- clean_data$campaign + clean_data$previous
head(clean_data)
```
This are the dataframe of clean_data after added three more columns which is age_group, recency_level and total_contact_score.<br>

**Interactive Relationship of Analysis**<br>
**A chart comparing two categorical variables**
```{r}
count_data <- clean_data %>%
  count(education, marital)

plot_ly(count_data,
        x = ~marital,
        y = ~n,
        color = ~education,
        type = "bar") %>%
  layout(barmode = "group",
         title = "Education Level across Marital Status",
         xaxis = list(title = "Marital", tickangle = -45),
         yaxis = list(title = "Count"))
```
Barplot of education across marital status.<br>
It can be seen that the "married" marital status category contains the largest number of individuals overall.Within three marital status, individuals with a university degree are the most numerous. Education levels such as "illiterate" and "basic.6y" consistently represent very small proportions across all marital statuses.

**A chart comparing one numeric and one categorical**
```{r}
plot_ly(data = clean_data,
        x = ~job,
        y = ~age,
        type = "box",
        split = ~job,
        box = list(visible = TRUE),
        meanline = list(visible = TRUE),
        points = "all") %>%
  layout(title = "Age Distribution Across Job Levels",
         xaxis = list(title = "Job", tickangle = -45),
         yaxis = list(title = "Age"),
         showlegend = FALSE)
```
Boxplot of Age Distribution Across Job.<br>
Overall, retired individuals have the highest age range, with a median age of 59 and a wide spread. On the other hand, students are the youngest group, with a median age of 26. Jobs such as housemaid and management tend to have higher median ages which is 40-45 years old, In contrast, categories like admin, services, and technician show slightly lower medians. Outliers are present in nearly all job categories, particularly among retired groups.<br>

**A multi-dimensional chart**
```{r}
p<- ggplot(clean_data, aes(x = age, fill = contact)) +
  geom_histogram(bins = 20, alpha = 0.7, position = "dodge") +
  facet_wrap(~ housing, scales = "free_y") +
  labs(title = "Age by Contact Status across Housing Loan",
       x = "Age (years)", 
       y = "Count",
       fill = "Contact Status") +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

ggplotly(p)
```
Age Distribution by Contact Status across Housing.<br>
Cellular is the primary contact method across all age groups and housing loan outcomes, with the most active contact occurring in the 30-40 age range.

### Insight Storytelling
**Data Story Telling of Loan Types by Marital Status – Subscribed Clients Only**
```{r}
marital_housing_data <- clean_data %>%
  filter(y == "yes") %>%
  group_by(marital) %>%
  summarise(
    housing_rate = round(mean(housing == "yes") * 100, 1),
    personal_loan_rate = round(mean(loan == "yes") * 100, 1),
    .groups = "drop"
  )

housing_max <- max(marital_housing_data$housing_rate)
loan_max <- max(marital_housing_data$personal_loan_rate)

loan_data <- data.frame(
  marital = rep(marital_housing_data$marital, 2),
  rate = c(marital_housing_data$housing_rate,
           marital_housing_data$personal_loan_rate),
  loan_type = rep(c("Housing Loan", "Personal Loan"), each = nrow(marital_housing_data))
)

loan_data$color <- ifelse(
  (loan_data$loan_type == "Housing Loan" & loan_data$rate == housing_max) |
  (loan_data$loan_type == "Personal Loan" & loan_data$rate == loan_max),
  "#d62728",  
  ifelse(loan_data$loan_type == "Housing Loan", "#AFEEEE", "#2ca02c")
)

loan_data$x_pos <- ifelse(loan_data$loan_type == "Housing Loan", 0, 0.4)

p <- ggplot(loan_data, aes(x = marital, y = rate)) +
  geom_col(aes(fill = loan_type), 
           position = position_nudge(x = loan_data$x_pos), 
           alpha = 0.8, width = 0.4, color = loan_data$color) +  
  geom_text(aes(label = paste0(rate, "%")), 
            position = position_nudge(x = loan_data$x_pos), 
            vjust = -0.3, size = 4) +
  scale_fill_manual(values = c(
    "Housing Loan" = "#AFEEEE",  
    "Personal Loan" = "#2ca02c"  
  )) +
  labs(
    title = "Loan Types by Marital Status – Subscribed Clients Only",
    x = "Marital Status",
    y = "Loan Rate (%)",
    fill = "Loan Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


ggplotly(p)

```
**Findings**<br>
Among clients who subscribed to a term deposit, married individuals show the highest engagement in both housing and personal loans (53% and 17.1% respectively). Single clients are close behind in housing loans at 52.6%, surprisingly matching married clients, but have a higher personal loan rate than divorced individuals. Divorced clients have the lowest rates for both loan types, particularly personal loans (11.8%). While there's a clear correlation between marital status and loan ownership, it's important to note that this doesn't imply causation; other factors like education and job could be at play.

**Data Story Telling of Loan Types by Education Level–Subscribed Clients Only**
```{r}
campaign_data <- clean_data %>%
  group_by(education) %>%
  summarise(
    total = n(),
    subscribed = sum(y == "yes"),
    success_rate = round((subscribed / total) * 100, 2)
  ) %>%
  arrange(desc(success_rate)) %>%
  mutate(education = factor(education, levels = education))

colors <- ifelse(campaign_data$education == "illiterate", "#e74c3c", "#3498db")

p <- plot_ly(
  data = campaign_data,
  x = ~education,  
  y = ~success_rate,  
  type = 'bar',
  marker = list(color = colors),
  text = ~paste0(success_rate, "%"),
  textposition = 'outside'
) %>%
  layout(
    title = "Term Deposit Subscription Rate by Education Level",
    xaxis = list(title = "Education Level"),
    yaxis = list(title = "Conversion Rate (%)"),
    showlegend = FALSE
  )

p

```
**Findings**<br>
The central trend indicates that clients with higher education levels (such as university.degree and high.school) generally have higher subscription rates, with university graduates reaching a 13.23% conversion rate. In contrast, those with basic education levels, especially basic.9y and basic.6y, display the lowest responsiveness, suggesting limited impact from the campaign in those segments.The most unusual segment in the chart is the illiterate group, which shows a remarkably high conversion rate of 16.6%, the highest among all education levels. This stands out because the group represents a very small portion of the dataset, meaning the high rate could be statistically unreliable or misleading. The chart suggests a possible correlation between education level and subscription rate with higher education levels generally linked to better responsiveness.

**Recommendation**<br>
"Target married clients for bundled offers involving both housing and personal loans, as they show the highest engagement across both categories which is 53% for housing loans and 17.1% for personal loans."<br>
"Prioritize clients with university degrees or high school education in future campaigns, as they combine large group sizes with consistently high subscription rates (13.23% and 12.06% respectively)."
